% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={BY LAWS},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{BY LAWS}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\newpage{}

BY LAWS OF THE\\
VPI CAVE CLUB\\
STUDENT GROTTO OF THE NATIONAL SPELEOLOGICAL SOCIETY\\
ADOPTED September 27, 2024

\hypertarget{safety-code}{%
\section{Safety Code}\label{safety-code}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  General Safety:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Never Cave alone. It is recommended that at least three people be on
    any caving trip; at least one full or associate member on horizontal
    trips and at least two full or associate members on vertical trips.
  \item
    Everyone participating in a trip should be physically and mentally
    capable of caving on that type of trip. Mind altering substances
    will not be used immediately prior to or during a cave trip.
  \item
    No one should exert themselves or be forced to exert themselves
    beyond their limit of endurance.
  \item
    If all lights fail, wait in place.
  \item
    Avoid jumping.
  \item
    All equipment should be in working order and pass inspection before
    entering the cave.
  \end{enumerate}
\item
  Equipment

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    General equipment per person.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Hard hats with suspension, chinstraps, and a mounted light source
      are mandatory while caving.
    \item
      At least three (3) sources of light, each providing sufficient
      light to exit the cave under reasonable circumstances.
    \item
      Supplies for light sources.
    \item
      Heat source.
    \item
      Trash bag or space blanket.
    \item
      Energy ration (candy bar, etc.)
    \item
      Gloves, heavy boots or shoes, heavy clothing, or cave specific
      clothing are advisable.
    \end{enumerate}
  \item
    Appropriate type and quantity of equipment shall be taken on all
    caving trips.
  \end{enumerate}
\item
  Procedure

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Always inform someone of the whereabouts of the caving trip, the
    expected time of return and appropriate phone numbers.
  \item
    Observe surroundings: respect landowners' wishes and be considerate
    of their property.
  \item
    If possible, consult someone who is familiar with the cave.
  \item
    All rigging is to be checked by a full or associate member. Ropes
    should be protected from fraying. All group members should be
    satisfied with rigging before it is used.
  \item
    Never leave a prospective member or other inexperienced caver alone.
  \item
    A belay should be used on every person on a ladder. No belay shall
    be used when ascending rope. Prospective members or other
    inexperienced cavers must accept a belay when requested to by a full
    or associate member. Anyone desiring a belay shall have one. The
    first person to descend a rope shall ensure that those following
    have a belay. The order of ascent and descent shall be determined by
    a full or associate member on the trip.
  \item
    An appropriate loop knot should be tied in the lower end of any rope
    used for rappelling.
  \end{enumerate}
\end{enumerate}

\hypertarget{membership}{%
\section{Membership}\label{membership}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Qualifiers:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    The qualifier ``inactive'' applied to any of the membership
    categories shall denote that said individual has not paid current
    dues and therefore retains on the privilege of attending and
    addressing regular meetings. payment of dues shall result in
    automatic reinstatement of the membership previously held and all
    privileges thereof.
  \item
    The qualifier ``life'' shall be applied to any full, associate, or
    prospective member of the VPI Cave Club upon receipt of life
    membership dues. Life members shall have membership privileges for
    which they qualify, but are exempt from yearly dues.
  \end{enumerate}
\item
  Training Program:\\
  The VPI Cave Club training program shall develop educated cavers
  consistent with Article II of the Constitution. The training program
  shall develop the basic skills to be safe and self-reliant
  underground. To be qualified as a full or associate member of the
  club, a prospective member must:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Read and understand the constitution and by-laws of the Cave Club.
  \item
    Remain a prospective member for at least 10 weeks during which time
    he/she spends 40 hours underground on at least 6 trips on which at
    least one full or associate member if present.
  \item
    Demonstrate elementary climbing skills in a cave.
  \item
    Demonstrate a working knowledge of belaying methods.
  \item
    Be able to tie a seat harness with webbing.
  \item
    Demonstrate a working knowledge of changeover from descending to
    ascending and ascending to descending.
  \item
    Demonstrate a working knowledge of passing a rebelay while ascending
    and descending.
  \item
    Demonstrate a working knowledge of rappelling and climbing rope in a
    cave.
  \item
    Demonstrate how to rig and pad a rope correctly.
  \item
    Know how to tie (demonstrate the lock off and back off where
    appropriate) and explain the uses of: bowline, bowline on a bight,
    bowline on a coil, figure 8, figure 8 on a bight, double figure 8,
    alpine butterfly, square knot, overhand knot, Münter hitch (and
    lockoff), water knot, helical, prusik, double fisherman's bend.
  \item
    Complete satisfactory a comprehensive general information quiz.
  \item
    Be endorsed by a full or associate member in good standing.
  \item
    Submit to the editor a suitable article for publication in the Tech
    Troglodyte allowing reasonable time for review and revision.
  \item
    Be approved by a 2/3 majority vote of the full membership.
  \end{enumerate}
\end{enumerate}

\hypertarget{disciplinary-action}{%
\section{Disciplinary Action}\label{disciplinary-action}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Any prospective, associate, or full member who has allegedly committed
  an infraction of the VPI Cave Club Constitution, or by-laws, or upon
  recommendation of the Safety and Techniques Committee, may face
  possible disciplinary action in the form of a reprimand, suspension,
  or expulsion, as defined below.
\item
  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Levels of disciplinary action:

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Reprimand: A reprimand shall be the lowest level of disciplinary
      action and shall be noted in the minutes.
    \item
      Suspension: A suspension is the intermediate level of disciplinary
      action. A suspension shall be reduction of an associate or full
      member's rights to those equivalent to prospective member status.
      A prospective member cannot complete any membership requirements
      while on suspension. The time period for a suspension shall be
      fifteen (15) weeks after which time the previous membership status
      is reinstated.
    \item
      Expulsion: An expulsion is the highest and most severe level of
      disciplinary action. An expelled person is ejected from the club
      and club functions and current dues are returned. Previously
      completed membership requirements are null and void. A person
      expelled from the club shall not be allowed to restart the
      training program or regain any level of membership in the club.
    \end{enumerate}
  \item
    Procedure:

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      A member of any class may bring a complaint for disciplinary
      action to the president.
    \item
      The president shall convene the Executive Committee to evaluate
      the complaint.
    \item
      Upon review the Executive Committee shall bring the complaint to
      the club, within two regular meetings of the receipt of the
      complaint, with a recommendation for action.
    \item
      The Executive Committee's recommendation shall constitute a
      motion.
    \item
      If the motion is to dismiss the complain, upon a ``second'' it may
      be voted on at the current meeting and requires 2/3 majority vote
      to pass.
    \item
      If the motion is for a level of disciplinary action, as defined
      above, upon a ``second'' the motion will be tabled until the next
      regular meeting.
    \item
      The accused shall be given notice of the tabled motion at least
      three (3) days in advance of the meeting in which the final vote
      is to be taken and shall be given ample opportunity to present a
      defense before the club.
    \item
      At the meeting following the tabled motion a discussion on the
      accused's infractions will occur. At the close of the discussion a
      secret ballot shall be taken to determine whether or not action
      will be taken on the disciplinary measure.
    \item
      A vote to take action must be passed by 2/3 majority of full
      members.
    \end{enumerate}
  \end{enumerate}
\end{enumerate}

\hypertarget{officers}{%
\section{Officers}\label{officers}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The officers of the club shall be elected by ballot from candidates
  nominated from the floor. The order of the election shall be
  president, vice-president, treasurer and secretary. To be nominated, a
  member must meet the qualification of Article IV, Section 2 of the
  constitution. A separate ballot shall be voted for each office. A
  simple majority of the votes cast shall be necessary for election. In
  case there is no majority, the candidate with the lowest number of
  votes shall be dropped and a re-vote for the remaining candidates
  shall be taken. The election shall be held at the tenth regular
  meeting of the spring semester. A 2/3 vote of full members shall be
  necessary to invalidate the election.
\item
  If the president and vice-president are absent during the summer, the
  president shall appoint a president pro tem from among the full
  members available.
\item
  in event of the resignation or release from office of an officer, a
  special election shall be held at the next regular meeting to elect a
  member to the office for the completion of the term of office.
\item
  The duties of the president shall be:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    To preside at all meetings and call special meetings.
  \item
    To act as an ex-officio member of all committees.
  \item
    To represent the club officially in all matters except those which
    require the cooperation of the treasurer.
  \item
    To make appointments to special committees unless provision is made
    otherwise in the resolution establishing the committee with the
    approval of the majority of the committee members.
  \item
    To make appointments to standing committees and to appoint committee
    chairmen and to remove appointees.
  \item
    To make arrangements for the meeting room.
  \item
    To appoint an editor for club publications.
  \item
    To appoint a club parliamentarian.
  \end{enumerate}
\item
  The duties of the Vice-President shall be:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    To assume the president's duties during the president's absence.
  \item
    To assist the president as the president shall require.
  \item
    To countersign as authorizing officer, vouchers fo the expenditures
    that have been authorized by the club.
  \item
    To be an active participant in the safety committee as director of
    the new member training program.
  \end{enumerate}
\item
  The duties of the Treasurer shall be:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    To have charge of the finances of the club.
  \item
    To keep an accurate record of the finances of the club, including
    balancing records weekly.
  \item
    To collect any and all dues and fees which the club shall authorize.
  \item
    To give a complete report on the financial status of the club at
    every regular meeting.
  \item
    To keep a list of the club members and their status, and to make
    this available to all members, faculty advisors, and at all club
    meetings.
  \item
    To make disbursements when authorized by motions passed by the club
    and only when so authorized.
  \item
    To comply with regulations of the VPI Office of Student Affairs
    pertaining to finances.
  \item
    To allow no transactions to occur without proper receipts.
  \end{enumerate}
\item
  The duties of the Secretary shall be:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    To keep a record of all club proceedings of all meetings of the club
    and place a copy in the club files, within two academic weeks.
  \item
    To keep a phone list of the members of the club and to make this
    available to all members and faculty advisors.
  \item
    To conduct correspondence for the club as directed by the president.
  \item
    To notify the NSS of changes in officers and other pertinent
    information.
  \item
    To notify faculty advisors of proposed amendments to the
    Constitution or the By-Laws.
  \item
    To make note of proxy votes.
  \item
    To provide timely notifications to club members of issues as
    specified by the constitution or by-laws.
  \end{enumerate}
\item
  Impeachment

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    An officer may be impeached for specific and willful infractions of
    the Constitution, the By-Laws, or the policies of the club.
  \item
    Process:

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      A motion to impeach stating specific infractions shall be made by
      a member, committee, or other officer.
    \item
      The motion shall be automatically tabled upon a ``second'' to the
      next regular meeting.
    \item
      The officer shall be given notice of the motion at least three (3)
      days in advance of the meeting in which the vote is to be taken,
      and be given ample opportunity to present a defense before the
      club.
    \item
      Final action shall be by secret written ballot and will require a
      2/3 majority vote of the full membership for removal from office.
    \end{enumerate}
  \end{enumerate}
\item
  Officers who are full members at the time of election shall assume
  their duties at the close of elections. Officers who are associate
  members at the time of election shall assume their duties at the first
  meeting of the fall semester. All officers shall serve until the
  elections of the following spring. If any office is left vacant by the
  election of an associate member, the president shall appoint an
  officer pro tem from among the full members available.
\end{enumerate}

\hypertarget{advisor}{%
\section{Advisor}\label{advisor}}

If desired, any person may be elected as an advisor to the club by a
simple majority of votes from full membership.

\hypertarget{meeting-and-quorum}{%
\section{Meeting and Quorum}\label{meeting-and-quorum}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  To constitute a meeting for the transaction of business, a quorum
  shall constitute one half (1/2) of the full membership either in
  person or represented by proxy.

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    The secretary shall be notified of each proxy before a vote. Blanket
    or continuous proxies shall not be held.
  \item
    meetings shall be held Fridays at 7:00 p.m. during the academic
    year.
  \item
    Meetings shall comply with Robert's Rules of Order.
  \end{enumerate}
\end{enumerate}

\hypertarget{dues}{%
\section{Dues}\label{dues}}

Dues for prospective members, full members and associate members shall
be ten dollars (\$10.00) per semester. Life membership shall be one
hundred and thirty dollars (\$130.00), payable in installments, to be
paid in full within one year of the first installment.

\hypertarget{committees}{%
\section{Committees}\label{committees}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  It shall be the duty of the Safety and Techniques Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Supervise the new member training program.
  \item
    Encourage safe caving practices and recommend disciplinary action
    where necessary.
  \item
    Present periodic programs on current safety practices.
  \item
    Submit a written report to the president upon request.
  \item
    Oversee maintenance and inventory of club rescue gear.
  \item
    The Safety and Techniques Committee shall consist of the safety
    chairman, the president, the vice-president, former safety chairmen,
    former presidents, and vice presidents, and anyone that the
    president or safety chairman deems necessary.
  \end{enumerate}
\item
  It shall be the duty of the Publicity Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Communicate with incoming freshman and the student body on the
    availability and activities of the club.
  \item
    Submit articles to the school newspaper, etc., on the club's
    activities.
  \item
    Publicize any activities which may be open to other groups.
  \item
    Submit to the files such material that may be printed by or about
    the club.
  \item
    Submit a written report to the president upon request.
  \end{enumerate}
\item
  It shall be the duty of the Conservation Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    The chairman of the conservation committee shall be a member of the
    National Conservation Committee of the NSS.
  \item
    Encourage projects and better practices in cave conservation.
  \item
    Instruct new members on the principles of cave conservation.
  \item
    To bring to the attention of the club any violation of the club's
    policy of conservation.
  \item
    To scrutinize the mailbag for any material related to cave
    conservation.
  \item
    Submit a written report to the president upon request.
  \end{enumerate}
\item
  It shall be the duty of the Project and Program Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Plan and coordinate club projects, special activities, and social
    functions
  \item
    Plan and make arrangements for programs to be presented periodically
    at the regular meetings.
  \item
    Submit a written report to the president upon request.
  \end{enumerate}
\item
  It shall be the duty of the Supplies Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Buy equipment for the club store for sale to the club members and
    prospective members out of a special revolving fund established for
    this purpose.
  \item
    Make equipment available for sale at all meetings to all dues paying
    individuals.
  \item
    The fund shall be increased by semesterly appropriations until such
    time that it becomes self-sustaining. The fund shall be allowed to
    draw upon the club treasury in the form of loans when additional
    funds become necessary.
  \item
    Keep an accurate and up-to-date record of the finances of the club
    store, including giving a financial status report within one week of
    a request from the treasurer, and submitting a complete report of
    the store's activities to the treasurer at the end of the fall and
    spring semesters.
  \end{enumerate}
\item
  It shall be the duty of the Executive Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Preside over disciplinary proceedings.
  \item
    Act for the club in emergency situations when it is not feasible for
    the club to meet.
  \item
    The Executive Committee shall consist of the officers, the head of
    the Safety and Techniques Committee and other person(s) appointed by
    the president.
  \end{enumerate}
\item
  It shall be the duty of the Equipment Committee to:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Maintain an inventory of Club equipment.
  \item
    Maintain the storage and dispersal access of Club equipment.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Club equipment shall be checked-out by dues-paid members as
      restricted in Article III.
    \item
      Club equipment shall only be used to support Article II.
    \end{enumerate}
  \item
    Establish an equipment check-out policy to include how long
    equipment may be checked-out and the required condition of equipment
    upon return.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Gear checked-out that does not conform to this policy, is grounds
      for a 21-day suspension from the privilege of accessing the Club
      equipment cache.
    \item
      Club equipment not checked-in, or checked-in damaged from
      negligence, shall be replaced by the Member who checked out the
      equipment.
    \end{enumerate}
  \item
    Maintain Club equipment in a safe condition, and quarantine
    equipment that is not in safe condition.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Quarantined equipment may be sold, gifted, repaired, or destroyed
      at the discretion of the Equipment Committee. All gear sale
      proceeds will return to the Club General Fund.
    \end{enumerate}
  \item
    Maintain Club identification markings on all non-infrastructure Club
    equipment.
  \item
    Identify equipment capability gaps for common Club activities.
  \item
    Recommend the acquisition of new equipment when necessary.
  \end{enumerate}
\end{enumerate}

\hypertarget{club-files}{%
\section{Club Files}\label{club-files}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Maintenance: The president shall appoint a person to keep and maintain
  the files.
\item
  Substance: The club files shall consist of cave maps, trip reports,
  meeting minutes, cave related publications, membership related
  material, and historical material.
\item
  The file chairman shall submit a written report to the president upon
  request.
\end{enumerate}

\hypertarget{club-trips}{%
\section{Club Trips}\label{club-trips}}

Club trips are caving trips which are recognized by the VPI Cave Club
and which follow the spirit and intent of its constitution and by-laws.
A caving trip is recognized if it has at least one full or associate
member on it and it is properly signed out. (See By-Laws, Safety Code,
Section 3.A for definition of a properly signed out trip.)

\hypertarget{club-savings}{%
\section{Club Savings}\label{club-savings}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The club shall establish a savings fund to be placed in an interest
  drawing savings account.
\item
  The savings fund shall be drawn upon only for investments of long
  range importance to the club.
\item
  The savings fund shall be increased by semesterly appropriations when
  possible.
\end{enumerate}

\hypertarget{club-publication}{%
\section{Club Publication}\label{club-publication}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The grotto shall publish a periodic journal to advertise grotto
  projects, trips, and other activities. It shall further contain
  material of scientific and technical interest to the caving world.
\item
  The title of this publication shall be: The Tech Troglodyte.
\item
  The format and journalistic style shall be set by the editor.
\item
  The circulation of this publication shall be to all members and
  prospective members who have paid current dues. In addition, the NSS
  library and all other grottos which offer their publication in
  exchange shall receive this publication. An exchange editor will be
  appointed by the president for these duties, and he/she will cooperate
  with the mailbag person to keep an updated exchange file.
\end{enumerate}

\end{document}
