% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={CONSTITUTION},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{CONSTITUTION}
\author{}
\date{}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\newpage{}

CONSTITUTION OF THE\\
VPI CAVE CLUB\\
STUDENT GROTTO OF THE NATIONAL SPELEOLOGICAL SOCIETY\\
ADOPTED April 26, 2024

\hypertarget{article-i-name}{%
\section{Article I: Name}\label{article-i-name}}

The name of this organization shall be the VPI Cave Club, Student Grotto
of the National Speleological Society (VPI is an acronym for Virginia
Polytechnic Institute.)

\hypertarget{article-ii-purpose}{%
\section{Article II: Purpose}\label{article-ii-purpose}}

The purpose of this club shall be to promote interest in, and to advance
the science of speleology; to promote conservation of caves and safety
in their exploration; and to encourage fellowship among those interested
in caving.

\hypertarget{article-iii-membership}{%
\section{Article III: Membership}\label{article-iii-membership}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  There shall be three classes of dues paying members: full, associate,
  and prospective. There shall be two qualifiers, inactive and life, as
  described in the By-Laws.

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Full membership shall be restricted to students, faculty, or staff
    of VPI\&SU who have completed membership requirements as stated in
    the By-Laws. Full members shall have access to club records and
    files and may use club equipment.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Life members are considered to be associate (defined below) until
      they have submited written notice to the Secretary or Treasurer
      each semester. Non-life full members are automatically registered
      to vote when their dues are paid.
    \item
      Any full members may choose associate membership. Full membership
      may be reinstated upon written notification of the club treasurer
      during the first two meetings of any subsequent semester.
    \end{enumerate}
  \item
    Associate membership shall be reserved for those who have completed
    requirements for full membership but are not students, faculty or
    staff of VPI\&SU. Associate members have all privileges of full
    membership except for the holding of an elected office and voting.
    Full and associate members shall have access to club records and
    files and may use club equipment.

    \begin{enumerate}
    \def\labelenumiii{\roman{enumiii}.}
    \tightlist
    \item
      Faculty or staff who are full members may for any semester for
      which dues have been paid choose associate membership. Full
      membership may be reinstated upon written notification of the club
      treasurer during the first two weeks of any subsequent semester.
    \item
      Students who are full members and who anticipate being absent from
      campus for a semester may choose associate membership for that
      semester.
    \end{enumerate}
  \item
    Any individual may be invited verbally by the president and
    vice-president to become a prospective member upon recommendation of
    a full or associate member. The prospective member must be willing
    to uphold the purposes and policies of the club. A prospective
    member may attend and address meetings, but may not vote or hold
    elected office.
  \end{enumerate}
\item
  Honorary members may be recognized at the discretion of the club for
  their outstanding achievements or contributions to the club or the
  interests of the club. Honorary membership is a ceremonial title and
  holes no bearing on actual membership status.
\item
  Dues shall be paid in a timely manner in an amount as stated in the
  by-laws.
\item
  Any prospective, associate, or full member may face disciplinary
  action, including possible loss of membership privileges, as defined
  in the By-Laws.
\end{enumerate}

\hypertarget{article-iv-officers}{%
\section{Article IV: Officers}\label{article-iv-officers}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The elected officers of this club shall be: president, vice-president,
  secretary, and treasurer. Election procedures are defined in the
  By-Laws.
\item
  All elected officers must be enrolled as full-time students at VPI\&SU
  throughout the duration of their term. The President and Vice
  President and Treasurer must be full or associate members at the time
  of the election. The Secretary can be a full, associate or prospective
  member. If an officer is not an NSS member, membership must be applied
  for, within one month of the election.
\item
  In the event of extenuating circumstances, including but not limited
  to events preventing meetings, caving, training of new members, or
  other club activities, the voting membership of the club shall have
  the option to declare a state of emergency by 2/3rds majority vote.
  The purpose of the state of emergency will be to allow the club to
  temporarily suspend portions of the bylaws that prevent normal
  operations of the club because of the conditions of the emergency.Upon
  the end of the state of emergency a special election will take place
  at the next regular meeting in order to fill cave club officer
  positions with traditionally eligible members. Should the vote to end
  the state of emergency fall at a time when meetings are not happening,
  the vote will occur at the next meeting of the club, and the state of
  emergency will extend until it is voted upon at that said meeting.
  During a state of emergency the following policies will go into
  effect:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    Any full, student member may hold office regardless of part-time or
    full-time status.
  \item
    In the event that there are not enough willing student members to
    hold office, the club may elect non-student associate and full
    members, as defined in Article III, Section 1.b to take on the
    responsibilities of the Vice President and/or Secretary as acting
    officers
  \item
    A state of emergency has no effect on a member's voting status,
    except,
  \item
    In the event that fewer than 3 members are eligible to vote,
    associate members will be allowed to vote to approve new members.
  \item
    During a state of emergency the items herein supersede any and all
    conflicting bylaws or directives
  \end{enumerate}
\end{enumerate}

\hypertarget{article-v-committees}{%
\section{Article V: Committees}\label{article-v-committees}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Standing committees shall be: the Equipment Committee, Safety and
  Techniques Committee, Publicity Committee, Project and Program
  Committee, Conservation Committee, Supplies Committee and Executive
  Committee.
\item
  Special and ad hoc committees may be established by the President or
  by action of the club by simple majority vote.
\end{enumerate}

\hypertarget{article-vi-amendments}{%
\section{Article VI: Amendments}\label{article-vi-amendments}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Amendments to this Constitution may be submitted by motion of any full
  member and then tabled after passing with a simple majority of votes
  from full membership.
\item
  An amendment must be voted on no later than three meetings subsequent
  to its proposal, and the full and associate members of the club must
  be specifically notified of the date of the vote and the nature of the
  amendment, in writing.
\item
  A 3/4 majority vote of the full membership shall be required for the
  adoption of any amendment.
\end{enumerate}

\hypertarget{article-vii-by-laws}{%
\section{Article VII: By-Laws}\label{article-vii-by-laws}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The club shall establish a set of By-Laws for the purpose of
  amplifying and explaining in detail certain sections of the
  Constitution.
\item
  The By-Laws may be amended as follows:

  \begin{enumerate}
  \def\labelenumii{\Alph{enumii}.}
  \tightlist
  \item
    A motion to amend the By-Laws may be introduced by any full or
    associate member.
  \item
    If the motion is ``seconded'' it is automatically tabled.
  \item
    The secretary must notify the members of the amendment in writing.
  \item
    Final action shall be taken at the next meeting by a 2/3 vote of the
    full membership.
  \end{enumerate}
\end{enumerate}

\hypertarget{article-viii-dissolution}{%
\section{Article VIII: Dissolution}\label{article-viii-dissolution}}

Pending a ¾ or greater full membership vote in favor of dissolving, all
assets remaining after meeting outstanding liabilities shall be assigned
to the National Speleological Society. However, if the named recipient
is not in existence or no longer is a qualified distributee, or
unwilling or unable to accept the distribution, the assets of this
organization shall be distributed to a fund, foundation, or corporation
organized and operated exclusively for the purposes specified in Section
501 (c) (7) of the Internal Revenue Code of 1954 (or corresponding
provision of any future U. S. Internal Revenue Law).

\end{document}
